---
title: "Logistic Regression and Monte Carlo Simulation"
author: | 
    | Fall 2022, MATH8050: Homework 5
    | **Ashwini Balasubramanian(C31547239), Section 001**
date: "Due October 13, 12:00 PM noon"
output: 
     pdf_document:
font-size: 12pts
urlcolor: blue
keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=TRUE, include=TRUE, results="hide"}
# load packages 

library(ggplot2)
library(stats)
library(ISLR2)

sessionInfo()
```
# 1a
```{r}
summary(Smarket)
fit.sm = glm(Direction~Lag1+Lag2+Lag3+Lag4+Lag5+Volume,data=Smarket,family=binomial(link="logit"))
summary(fit.sm)
coef(fit.sm)
```
# 1b

```{r}
func1 = function(X,y,threshold = 1e-10, max_iter = 100) {
p1 = function(X,beta)
{
    beta = as.vector(beta)
    return(exp(X%*%beta) / (1+ exp(X%*%beta)))
  }
beta = rep(0,ncol(X))
diff = 10000
iter_count = 0
while(diff > threshold ) 
{
p = as.vector(p1(X,beta)) 
W = diag(p*(1-p))
beta_change = solve(t(X)%*%W%*%X) %*% t(X)%*%(y - p)
beta = beta + beta_change
diff = sum(beta_change^2)
iter_count = iter_count + 1 
if(iter_count > max_iter) {
      stop("This isn't converging")
    }
}
  return(beta)
}
new_Smark<-Smarket
new_Smark<-transform(new_Smark,new_d=ifelse(Direction=="Up",1,0))
X.temp1<-Smarket$Lag1 
X.temp2<-Smarket$Lag2 
X.temp3<-Smarket$Lag3 
X.temp4<-Smarket$Lag4 
X.temp5<-Smarket$Lag5 
X.Volume<-Smarket$Volume
y <- as.numeric(new_Smark$new_d) 
n <- nrow(Smarket)
X<-cbind(c(rep(1,n)),X.temp1,X.temp2,X.temp3,X.temp4,X.temp5,X.Volume) 
colnames(X) <- c("Intercept","Lag1","Lag2","Lag3","Lag4","Lag5","Volume")
func1(X,y)
```
# 1c
```{r}
set.seed(12345)
sample_size = floor(0.85*nrow(Smarket))
train_ind = sample(seq_len(nrow(Smarket)),size = sample_size)
train =Smarket[train_ind,]
test=Smarket[-train_ind,]
fit.sm_train= glm(formula = Direction ~ Lag1+Lag2+Lag3+Lag4+Lag5+Volume,
              data=train,family = binomial)
beta<-as.data.frame(coef(fit.sm_train)) 
betas<-c(rep(0,7))
c=1
for (i in beta$`coef(fit.sm_train)`) {
betas[c]=i
c=c+1 }
p= function(x1,x2,x3,x4,x5,x6,betas,n) { 
c=1
val<-c(rep(n),0) 
ans<-c(rep(n),0) 
i=1
while ( i <=n){
    X2<-c(x1[c],x2[c],x3[c],x4[c],x5[c],x6[c])
    val[c]<-betas[1]+sum(X2*betas[2:7])
    ans[c]<-1/(1+exp(-val[c]))
    c=c+1
    i=i+1
}
  return (ans)
}
x1<-test$Lag1
x2<-test$Lag2
x3<-test$Lag3
x4<-test$Lag4
x5<-test$Lag5
x6<-test$Volume
n=nrow(test)
values_p1<-p(x1,x2,x3,x4,x5,x6,betas,n)
test_direction<-test$Direction
train_direction<-train$Direction
my.predict= function(test_direction,values_p){ test_data_direction<-c(rep(length(test_direction),0)) 
c=1
for (i in values_p)
{
if (i>0.5){ 
  test_data_direction[c]<-"Up"
} 
  else{
      test_data_direction[c]<-"Down"
    }
c=c+1
}
c=1
true_down=0
true_up=0
true_down_butup=0 
true_up_butdown=0
for (i in test_data_direction) {
if(test_direction[c] == "Down" && test_data_direction[c] =="Down") {
     true_down=true_down+1
  }
else if(test_direction[c] == "Up" && test_data_direction[c] == "Up"){
true_up=true_up+1
}
else if(test_direction[c] == "Up" && test_data_direction[c] == "Down"){ 
  true_up_butdown=true_up_butdown+1
}
else if(test_direction[c] == "Down" && test_data_direction[c] == "Up"){
      true_down_butup=true_down_butup+1
    }
c=c+1 
}
result<-data.frame(true_down=c(true_down),true_up=c(true_up),
                           true_down_butup=c(true_down_butup),
                           true_up_butdown=c(true_up_butdown))

return(result)
}
my.predict(test_direction,values_p1)
```

The false positive is 75 and false negative is 12 for the test data
```{r}
x1<-train$Lag1
x2<-train$Lag2
x3<-train$Lag3
x4<-train$Lag4
x5<-train$Lag5
x6<-train$Volume
n=nrow(train)
values_p<-p(x1,x2,x3,x4,x5,x6,betas,n)
my.predict(train_direction,values_p)
```
Therefore we can see that false positive and fasle negative are 75 and 430 respectively. We can verify that using predict funtion in r
```{r}
glm.probs <- predict(fit.sm_train,newdata = train, type="response")
glm.pred <- rep("Down", nrow(train))
glm.pred[glm.probs > 0.5] <- "Up"
glm.pred <- ifelse(glm.probs > 0.5,"Up","Down")
table(glm.pred, train$Direction)
```
```{r}
glm.probs <- predict(fit.sm_train,newdata = test, type="response")
glm.pred <- rep("Down", nrow(test))
glm.pred[glm.probs > 0.5] <- "Up" 
glm.pred <- ifelse(glm.probs > 0.5,"Up","Down")
table(glm.pred, test$Direction)
```

# 2a
The given equation can be written as:
\[ \int_{-\infty}^{\infty} exp(-x^4)=2\int_{0}^{\infty}exp(-x^4)
\]
Replacing $-x^4=t$ and thus $-t^{1/4}=x$
\[dx=\frac{1}{4}t^{-3/4}dt
\]
Rewriting the equation w.r.t t:
\[2\int_{0}^{\infty}exp(-x^4)=2/4\int_{0}^{\infty}t^{\frac{1}{4}-1}exp(-t)dt
\]
This is in the gamma function form,
\[\Gamma(z)=\int_{0}^{\infty}exp(-t)t^{z-1}dt
\]
where \[z= 1/4 \Gamma(1/4)=3.6356\]
therefore, \[0.5 \int_{0}^{\infty}t^{\frac{1}{4}-1}exp(-t)dt=\Gamma(1/4)0.5 = 1.8128
\]

# 2b
```{r}
h = function(x){return(2*exp(-(x^4))/dgamma(x,shape = 1.2,rate=0.95))}
val<- function(x) 2*exp(-(x^4)) 
integrate(val, 0, Inf)
```
```{r}
I = function(n, x){
hx = h(x)
hbar = mean(hx)
v = sum((hx-hbar)^2) / n^2 
L = hbar - 2*sqrt(v)
U = hbar + 2*sqrt(v)
return(list(hbar=hbar, v=v, L=L, U=U))
}
n=10^4
set.seed(29)
x = rgamma(10^4,shape=1.2,rate=0.95)
ans<-I(n, x)$hbar
c(mean(h(x)),var(h(x)))
```
As the above integration limits follow the limits of Gamma Distribution so we have taken the Gamma Distribution with shape and rate as 1.2 and 0.95.

# 2c
```{r}
set.seed(29)
w<-function(x) dgamma(x,shape=1.2,rate = 0.95)/dexp(x,2.2)
X<-rexp(10^5,2.2)
Y<-w(X)*h(X)
c(mean(Y),var(Y))
```
With the help of sampling we can see the variance of the value is reduced and also the precision of the value is increased at rate=2.2.
